### 2022-04-16 proposal/proposal.tex
Add proposal and papers.

\documentclass{amsart}

\usepackage{hyperref}

\title{COSC-373: Final Project Proposal}
\author{Lee Jiaen}
\author{Hyery Yoo}
\author{Alexander Lee}

\begin{document}

\maketitle

Recently, there has been a lot of hype about Bitcoin and other related
cryptocurrencies. We have therefore decided to investigate Bitcoin and its
blockchain technology for the final project. As a resource, we will read the
original Bitcoin paper by Satoshi Nakamoto
(\url{https://bitcoin.org/bitcoin.pdf}).

\end{document}